---
title: "Surface Urban Heat Island Analysis"
subtitle: "`r paste(params$city, '|', params$season, 'season', params$year)`"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
author: "SUHII Mapping Tool — Officina SCIFT"
format:
  html:
    theme:           cosmo
    toc:             true
    toc-depth:       3
    toc-title:       "Contents"
    toc-location:    left
    number-sections: true
    embed-resources: true    # single self-contained HTML file
    fig-width:       9
    fig-height:      6
    code-fold:       true
    code-summary:    "Show code"
    css:             report_style.css
params:
  city:          "Unknown"
  city_slug:     "unknown"
  season:        "warm"
  year:          2024
  start_date:    "2024-06-01"
  end_date:      "2024-08-31"
  max_cloud:     30
  output_dir:    "."
  koppen_class:     "Csa"
  altitude_band_height: 100
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

# Packages (already loaded by run.R via 00_packages.R)
library(terra)
library(leaflet)
library(leaflet.extras)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
library(scales)
library(tibble)

# ── Load all outputs ──────────────────────────────────────────────────────────
out  <- params$output_dir
slug <- params$city_slug
seas <- params$season
yr   <- params$year

f <- function(...) file.path(out, sprintf(...))

# Rasters
lst_mean    <- terra::rast(f("%s_%d_LST_MEAN.tif",             seas, yr))
anomaly     <- terra::rast(f("%s_%d_thermal_anomaly.tif",      seas, yr))
suhi        <- terra::rast(f("%s_%d_SUHI.tif",                 seas, yr))
classified  <- terra::rast(f("%s_%d_anomaly_classified.tif",   seas, yr))
priority    <- terra::rast(f("%s_%d_priority_map.tif",         seas, yr))

dist_path   <- f("%s_%d_distance_green_areas.tif", seas, yr)
has_dist    <- file.exists(dist_path)
if (has_dist) green_dist <- terra::rast(dist_path)

# Statistics
stats_path <- f("%s_%d_city_stats.csv", seas, yr)
stats      <- read.csv(stats_path)

# Classification labels and colours
CLASS_LABELS <- c(
  "Cool island",
  "Neutral",
  "Weak",
  "Moderate",
  "Strong",
  "Extreme"
)
CLASS_BREAKS <- c(-Inf, 0, 1, 2.5, 4.5, 6.5, Inf)
CLASS_COLOURS <- c("#4575b4","#ffffbf","#fee090","#f46d43","#d73027","#a50026")

# Helper: project to WGS84 for leaflet
# method = "near" is required for categorical/classified rasters — bilinear
# (the default) interpolates between class integers at every boundary,
# producing non-integer values that colorFactor() can't match -> transparent
# "holes" along every class transition.
to_wgs84 <- function(r, method = "bilinear") terra::project(r, "EPSG:4326", method = method)

# Helper: format numbers with consistent rounding
fmt1 <- function(x) formatC(round(x, 1), format = "f", digits = 1)
fmt2 <- function(x) formatC(round(x, 2), format = "f", digits = 2)
fmt0 <- function(x) formatC(round(x, 0), format = "f", digits = 0, big.mark = ",")

# Pre-format the green deficit text (avoids complex inline if/else in a table cell)
pct_green_txt <- if (is.na(stats$pct_urban_beyond_300m_green)) {
  "N/A"
} else {
  paste0(stats$pct_urban_beyond_300m_green, "%")
}
```

::: {.callout-note}
## Quick summary

**`r params$city`** — `r params$season` season **`r params$year`**
(`r params$start_date` → `r params$end_date`, cloud cover ≤ `r params$max_cloud`%)

| Indicator | Value |
|:----------|------:|
| Mean urban LST | **`r fmt1(stats$lst_mean_urban_C)` °C** |
| Mean rural LST | **`r fmt1(stats$lst_mean_rural_C)` °C** |
| Mean LST anomaly (SUHII) | **`r fmt1(stats$suhii_mean_C)` °C** |
| Max LST anomaly | **`r fmt1(stats$suhii_max_C)` °C** |
| Urban area beyond 300 m from green | **`r pct_green_txt`** |
| Landsat scenes used | **`r stats$n_landsat_scenes`** |

:::

---

# What is urban heat island?

Urban areas are generally warmer than their surrounding countryside. This
phenomenon — known as the **Urban Heat Island (UHI) effect** — arises because
cities replace vegetated surfaces with impervious materials (asphalt, concrete,
rooftops) that absorb more solar radiation, retain heat longer, and cannot cool
through evaporation the way plants do. The absence of shade trees, the density
of buildings that trap heat between facades, and the heat released by vehicles
and air conditioners all amplify the effect.

The consequences are real and measurable: higher energy consumption for cooling,
reduced air quality, greater health risk during heatwaves (especially for elderly
and low-income residents), and increased mortality. Studies estimate that UHI can
increase peak urban electricity demand by 2–4% for every 1°C of warming above a
threshold of 15–20°C [@akbari1995].

This report focuses on the **surface** expression of the UHI, quantified from
Landsat satellite imagery.

---

# What this analysis measures — and what it does not

::: {.callout-important}
## LST is not air temperature — please read this carefully

The analysis in this report is based on **Land Surface Temperature (LST)**: the
radiometric skin temperature of surfaces as seen from space by the Landsat
satellite's thermal infrared sensor. LST measures how hot the ground surface is
— the asphalt of a parking lot, a rooftop, a field of grass.

**LST is not the same as the air temperature you feel when you walk outside.**
Over dark impervious surfaces in summer, LST can be 20–40°C higher than
near-surface air temperature. The difference depends on the surface material,
wind speed, humidity, and time of day [@voogt2003].

When we report that a neighbourhood has an LST anomaly of +4°C relative to
surrounding rural areas, this does **not** mean that residents experience air
temperatures 4°C warmer. Empirical research suggests that a **2°C LST anomaly
corresponds roughly to ~1°C in near-surface air temperature** (T~air~) anomaly,
though this ratio is highly variable [@voogt2003; @imhoff2010].

Why do we use LST then? Because it is:

- **Spatially consistent**: derived from the same sensor across the whole city
  in a single overpass
- **Freely available** at 30 m resolution globally via Landsat
- **Scientifically established** as a robust proxy for SUHII in hundreds of
  peer-reviewed studies [@peng2012; @zhou2014]
- **Useful for comparison**: even if absolute values differ from T~air~, LST
  anomalies reliably identify the *relative* hot and cool spots

Think of this analysis as a **thermal fingerprint** of the city: it tells you
*where* the surface is hottest relative to the surrounding countryside, and by
how much. It is a starting point for deeper analysis and targeted intervention,
not a direct measure of outdoor thermal comfort.

:::

**When was the image acquired?**
Landsat's overpass time is approximately **10:00 local time** (descending node).
This is a morning measurement. The classical air-temperature UHI is typically
strongest at night (2–4 hours after sunset) and weakest around midday [@oke1982].
Surface UHI, by contrast, can be strongest in the early afternoon when solar
loading is maximum. The morning overpass captures a transitional state — not the
daily peak, but still informative for identifying persistently warm surfaces.

---

# Data and methodology

## Satellite data

Landsat Collection 2 Level 2 (C2L2) surface temperature products were downloaded
from **Microsoft Planetary Computer** (free, anonymous access via the STAC API).
C2L2 products have been atmospherically corrected and provide calibrated Land
Surface Temperature values. The scale factor applied to convert Digital Numbers
to Kelvin is: $LST [K] = DN \times 0.00341802 + 149.0$, then converted to
Celsius: $LST [°C] = LST [K] - 273.15$ [@usgs_c2l2].

Only **Tier 1** scenes (highest geometric and radiometric quality) were used.
Scenes with more than `r params$max_cloud`% cloud cover were excluded.
Individual scenes with more than 70% cloud/NA pixels over the study area were
discarded. A seasonal mean was computed pixel-wise across all valid scenes.

The **QA_PIXEL** band was used for per-pixel cloud masking:

- Landsat 8/9: valid pixel = QA_PIXEL == 21824 (clear land, no cloud,
  no cloud shadow, no snow)
- Landsat 5/7: valid pixel = QA_PIXEL == 5440

For Landsat 7 scenes acquired after May 2003 (scan-line corrector failure),
an 11×11 focal mean filter was applied to fill the characteristic data gaps
before mosaicking.

**`r stats$n_landsat_scenes`** valid scenes were used in this analysis
(season: `r params$start_date` → `r params$end_date`).

## Reference areas

The **urban area** was defined from OpenStreetMap features: artificial land use
(commercial, residential, industrial, etc.), amenity, tourism, leisure
(recreational), aeroway, and highway polygon features.

The **rural reference area** was defined from natural vegetation (OSM `natural`
tag: grassland, forest, wood, scrub, etc.), agricultural land use (farmland,
meadow, orchard, etc.), and urban green spaces (parks, gardens, nature
reserves). A 100 m buffer around the urban mask was excluded from the rural
reference to avoid edge contamination.

The **Surface UHI Intensity (SUHII)** is computed as:
$$\text{SUHII}(x,y) = \text{LST}(x,y) - \bar{\text{LST}}_\text{rural}$$

where $\bar{\text{LST}}_\text{rural}$ is the spatial mean of LST over all rural
reference pixels within the same elevation band. This definition follows
@peng2012 and @imhoff2010.

## Elevation stratification

To isolate the thermal signature of urbanisation from orographic temperature
gradients (temperature decreases ~6.5°C per 1000 m of altitude), the study
area was divided into **`r params$altitude_band_height` m elevation bands**.
The SUHII was computed independently within each band and the results were
merged into a seamless mosaic. This approach is particularly important for
cities on hills or with significant topographic relief.

Source: SRTM Global 1 arc-second DEM (~30 m), downloaded via the OpenTopography
API [@farr2007].

## Koppen-Geiger climate classification

The seasonal analysis window was derived from the **Koppen-Geiger climate
classification** of the study area (class: **`r params$koppen_class`**). This
ensures that the "warm season" corresponds to the climatically meaningful hot
period for the specific city, rather than using fixed calendar months that may
not be appropriate across different climate zones [@kottek2006].

---

# Results

## Seasonal mean Land Surface Temperature

```{r map-lst, fig.cap="Seasonal mean LST (°C). Warmer colours indicate higher surface temperatures."}
lst_wgs  <- to_wgs84(lst_mean)
lst_vals <- terra::values(lst_wgs, na.rm = TRUE)
lst_rng  <- range(lst_vals)   # full min/max: every pixel is guaranteed in-domain

pal_lst <- colorNumeric(
  palette = "Spectral",
  domain  = lst_rng,
  reverse = TRUE,
  na.color = "transparent"
)

leaflet(options = leafletOptions(minZoom = 8)) |>
  addProviderTiles("CartoDB.Positron", group = "Light") |>
  addProviderTiles("Esri.WorldImagery", group = "Satellite") |>
  addRasterImage(lst_wgs, colors = pal_lst, opacity = 0.80,
                 group = "Mean LST") |>
  addLegend("bottomright", pal = pal_lst,
            values = lst_rng,
            title  = "Mean LST (°C)",
            labFormat = labelFormat(suffix = "°C", digits = 0)) |>
  addLayersControl(
    baseGroups   = c("Light", "Satellite"),
    overlayGroups = "Mean LST",
    options = layersControlOptions(collapsed = FALSE)
  )
```

::: {.callout-note collapse="true"}
## How to read this map
The colours show the **average daytime surface temperature** (approximately
10:00 local time) across all cloud-free Landsat scenes during the
`r params$season` season of `r params$year`. Red and orange areas are the
hottest surfaces — typically dense urban cores, industrial zones, and large
parking areas. Blue and green areas are cooler — parks, water bodies, and
agricultural land on the urban fringe.
:::

```{r lst-stats}
data.frame(
  Statistic = c(
    "Mean LST — urban area",
    "Mean LST — rural reference",
    "Max LST — urban area",
    "Difference (urban − rural)"
  ),
  Value = c(
    paste0(fmt1(stats$lst_mean_urban_C), " °C"),
    paste0(fmt1(stats$lst_mean_rural_C), " °C"),
    paste0(fmt1(stats$lst_max_urban_C),  " °C"),
    paste0(fmt1(stats$lst_mean_urban_C - stats$lst_mean_rural_C), " °C ▲")
  )
) |>
  kable(caption = "LST summary statistics",
        col.names = c("Statistic", "Value"),
        align = c("l","r")) |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

## Thermal anomaly and SUHII classification

The map below shows the **LST anomaly**: the deviation of each pixel's surface
temperature from the mean of the rural reference areas, within the same
elevation band. Positive values indicate surfaces warmer than the rural
surroundings; negative values indicate cooler surfaces (urban cool islands,
typically parks or water bodies).

```{r map-classified, fig.cap="Surface Urban Heat Island Intensity classified into 6 levels (LST anomaly relative to rural reference)."}
cls_wgs <- to_wgs84(classified, method = "near")

pal_cls <- colorFactor(
  palette = CLASS_COLOURS,
  domain  = 1:6,
  na.color = "transparent"
)

leaflet(options = leafletOptions(minZoom = 8)) |>
  addProviderTiles("CartoDB.Positron", group = "Light") |>
  addProviderTiles("Esri.WorldImagery", group = "Satellite") |>
  addRasterImage(cls_wgs, colors = pal_cls, opacity = 0.80,
                 method = "ngb", group = "SUHII classes") |>
  addLegend("bottomright",
            colors = CLASS_COLOURS,
            labels = c(
              "Cool island (< 0°C)",
              "Neutral (0–1°C)",
              "Weak (1–2.5°C)",
              "Moderate (2.5–4.5°C)",
              "Strong (4.5–6.5°C)",
              "Extreme (> 6.5°C)"
            ),
            title  = "LST anomaly class",
            opacity = 0.9) |>
  addLayersControl(
    baseGroups   = c("Light", "Satellite"),
    overlayGroups = "SUHII classes",
    options = layersControlOptions(collapsed = FALSE)
  )
```

::: {.callout-important}
## What the classes mean

The six classes are based on the **LST anomaly** (surface temperature
difference between the urban pixel and the rural reference, in °C). These are
**surface temperature differences**, not air temperature differences.
Thresholds follow @stewart2012 and @chen2022 and are expressed as:

| Class | LST anomaly | Interpretation |
|:------|:-----------:|:---------------|
| Cool island | < 0°C | Surface cooler than rural (parks, water) |
| Neutral | 0–1°C | No significant surface heat island |
| Weak | 1–2.5°C | Detectable surface heat island |
| Moderate | 2.5–4.5°C | Significant surface heat island |
| Strong | 4.5–6.5°C | Intense surface heat island |
| Extreme | > 6.5°C | Extreme hotspot; potential health risk zone |

The global average daytime SUHII across 419 major cities is **1.5 ± 1.2°C**
[@peng2012]. Values above 2.5°C indicate an above-average urban heat effect
for this city relative to its rural surroundings.

**Reminder**: to estimate the *air temperature* anomaly a resident might
perceive, divide the LST anomaly by approximately 2 [@voogt2003]. A "Strong"
class (4.5–6.5°C LST) corresponds roughly to 2–3°C warmer air temperatures
— which is significant for health, especially during heatwaves.
:::

```{r bar-classes, fig.cap="Percentage of urban area in each SUHII class."}
class_df <- data.frame(
  class = factor(CLASS_LABELS, levels = rev(CLASS_LABELS)),
  pct   = c(
    stats$pct_urban_cool_island,
    stats$pct_urban_neutral,
    stats$pct_urban_weak,
    stats$pct_urban_moderate,
    stats$pct_urban_strong,
    stats$pct_urban_extreme
  ),
  colour = CLASS_COLOURS
)

ggplot(class_df, aes(x = class, y = pct, fill = class)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = paste0(round(pct, 1), "%")),
            hjust = -0.1, size = 3.5, colour = "grey30") +
  scale_fill_manual(values = setNames(rev(CLASS_COLOURS), rev(CLASS_LABELS))) +
  scale_y_continuous(
    limits = c(0, max(class_df$pct, na.rm = TRUE) * 1.2),
    labels = function(x) paste0(x, "%")
  ) +
  coord_flip() +
  labs(
    x = NULL, y = "% of urban area",
    title = "Distribution of SUHII classes across urban area",
    caption = paste0(
      "LST anomaly classes (surface temperature difference vs rural reference).\n",
      "Source: Landsat C2L2, ", params$start_date, " – ", params$end_date
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.y = element_blank(),
    plot.caption        = element_text(size = 9, colour = "grey50")
  )
```

## SUHII index map

The SUHII index is a **normalised version** of the LST anomaly, rescaled from
0 (coolest point in the study area) to 1 (hottest point). It is useful for
identifying spatial patterns and hotspots at a glance, without reference to
absolute temperature values.

$$\text{SUHII index} = \frac{\text{LST} - \text{LST}_{min}}{\text{LST}_{max} - \text{LST}_{min}}$$

```{r map-suhi, fig.cap="SUHII normalised index (0 = coolest, 1 = hottest). Useful for identifying spatial hotspots."}
suhi_wgs  <- to_wgs84(suhi)
suhi_vals <- terra::values(suhi_wgs, na.rm = TRUE)

pal_suhi <- colorNumeric(
  palette  = "Spectral",
  domain   = c(0, 1),
  reverse  = TRUE,
  na.color = "transparent"
)

leaflet(options = leafletOptions(minZoom = 8)) |>
  addProviderTiles("CartoDB.Positron", group = "Light") |>
  addProviderTiles("Esri.WorldImagery", group = "Satellite") |>
  addRasterImage(suhi_wgs, colors = pal_suhi, opacity = 0.80,
                 group = "SUHII index") |>
  addLegend("bottomright", pal = pal_suhi,
            values = c(0, 1),
            title  = "SUHII index",
            labFormat = labelFormat(digits = 1)) |>
  addLayersControl(
    baseGroups    = c("Light", "Satellite"),
    overlayGroups = "SUHII index",
    options = layersControlOptions(collapsed = FALSE)
  )
```

## Distance from green areas — the 3-30-300 rule

The **3-30-300 rule** was introduced by urban forester Cecil Konijnendijk
[-@konijnendijk2022] as an evidence-based guideline for equitable urban greening:

> - **3 trees** visible from every home, school, or workplace
> - **30%** tree canopy cover in every neighbourhood
> - **300 m** maximum distance to the nearest park or quality green space

This analysis addresses the **300 m component**: areas shown on the map are
those where the nearest green space is more than 300 m away, indicating a
**green access deficit**. Research consistently links poor access to green
spaces with worse mental health outcomes, higher heat stress, and greater
vulnerability during heatwaves [@konijnendijk2022; @gascon2015].

```{r map-green, eval=has_dist, fig.cap="Areas beyond 300 m from any green space (3-30-300 rule). Grey = within 300 m (compliant). Coloured = green access deficit."}
dist_wgs  <- to_wgs84(green_dist)
dist_vals <- terra::values(dist_wgs, na.rm = TRUE)
dist_max  <- max(dist_vals, na.rm = TRUE)

pal_dist <- colorNumeric(
  palette  = "YlOrRd",
  domain   = c(300, dist_max),
  na.color = "transparent"
)

leaflet(options = leafletOptions(minZoom = 8)) |>
  addProviderTiles("CartoDB.Positron", group = "Light") |>
  addProviderTiles("Esri.WorldImagery", group = "Satellite") |>
  addRasterImage(dist_wgs, colors = pal_dist, opacity = 0.80,
                 group = "Green deficit") |>
  addLegend("bottomright", pal = pal_dist,
            values   = c(300, dist_max),
            title    = "Distance to green (m)",
            labFormat = labelFormat(suffix = " m", digits = 0)) |>
  addLayersControl(
    baseGroups    = c("Light", "Satellite"),
    overlayGroups = "Green deficit",
    options = layersControlOptions(collapsed = FALSE)
  )
```

```{r green-stat, eval=has_dist}
if (!is.na(stats$pct_urban_beyond_300m_green)) {
  cat(sprintf(
    "\n**%s%%** of the urban area is more than 300 m from the nearest green or\n",
    stats$pct_urban_beyond_300m_green
  ))
  cat("rural reference area — these residents have a green access deficit\n")
  cat("according to the 3-30-300 rule [@konijnendijk2022].\n")
}
```

## Where to act — Intervention priority map

The map below identifies **where green infrastructure investment would most
reduce heat exposure**. The priority index combines three factors:

| Component | Weight | Rationale |
|:----------|:------:|:----------|
| LST anomaly (normalised) | **50%** | Primary driver of heat stress |
| Green access deficit (normalised) | **30%** | Main lever for UHI mitigation |
| Urban land cover | **20%** | Restricts priority to actionable areas |

High-priority areas (red) are **hot, poorly served by green infrastructure, and
urban** — exactly where planting trees, creating parks, or installing green
roofs would have the greatest combined benefit for cooling and accessibility.

The weights are informed by @maragkogiannis2024 and @morabito2015, who found
that thermal exposure and green deficit are the dominant drivers of heat
vulnerability in urban areas. **The weights are explicit and can be adjusted
by the user in module 08 (`W_THERMAL`, `W_GREEN`, `W_URBAN`)** to reflect
local priorities.

```{r map-priority, fig.cap="Intervention priority index (0 = low priority, 1 = highest priority). Only urban pixels are shown."}
prio_wgs <- to_wgs84(priority)

pal_prio <- colorNumeric(
  palette  = c("#ffffcc","#feb24c","#f03b20","#bd0026"),
  domain   = c(0, 1),
  na.color = "transparent"
)

leaflet(options = leafletOptions(minZoom = 8)) |>
  addProviderTiles("CartoDB.Positron", group = "Light") |>
  addProviderTiles("Esri.WorldImagery", group = "Satellite") |>
  addRasterImage(prio_wgs, colors = pal_prio, opacity = 0.85,
                 group = "Priority") |>
  addLegend("bottomright", pal = pal_prio,
            values  = c(0, 1),
            title   = "Priority index",
            labFormat = labelFormat(digits = 2)) |>
  addLayersControl(
    baseGroups    = c("Light", "Satellite"),
    overlayGroups = "Priority",
    options = layersControlOptions(collapsed = FALSE)
  )
```

---

# What can be done?

The science of UHI mitigation is well-established. The most effective and
co-beneficial strategies are:

**Green infrastructure** — Trees are the single most effective urban cooling
intervention. Street trees reduce surrounding air temperatures by 2–8°C through
shading and evapotranspiration [@bowler2010]. Expanding parks, green roofs,
and green corridors in high-priority areas addresses both the heat island and
the green deficit identified by the 3-30-300 rule.

**Cool/reflective surfaces** — High-albedo road and roof surfaces reduce solar
absorption. Combined with green roofs, hybrid interventions can reduce LST by
2–3°C in tropical and Mediterranean cities [@santamouris2014].

**Permeable pavements** — Reduce surface temperatures by enabling evaporation
and improving urban water management.

**Urban planning** — Street orientation and canyon geometry influence shadow
patterns and wind flow. Compact high-rise areas can trap heat between buildings,
while well-designed open spaces create cooling corridors.

**Who should act?**

- **City administrations**: prioritise the high-priority areas identified above
  for tree-planting programmes, park creation, and building regulations
  requiring green roofs.
- **Neighbourhood associations**: advocate for street trees and small pocket
  parks in heat-exposed areas.
- **Developers**: incorporate green infrastructure from the design stage.
- **Citizens**: plant trees in private gardens and courtyards; advocate for
  shaded public spaces.

---

# Limitations and caveats

- **LST ≠ T~air~**: as explained at the top of this report, this analysis measures
  surface temperature, not the air temperature experienced by people. For
  health risk assessment, additional data (population density, building
  typology, social vulnerability) should be integrated.

- **Single season**: one season's data captures the thermal pattern for that
  specific year. Anomalously hot or cool years may affect results. Multi-year
  averages are more robust.

- **30 m resolution**: Landsat's pixel size is 30 m × 30 m. Features smaller
  than this (a single tree, a small courtyard) are not resolved.

- **Morning overpass**: the ~10:00 acquisition time does not capture the
  nocturnal peak of the air-temperature UHI, which typically occurs 2–4 hours
  after sunset.

- **OSM data completeness**: the urban and rural masks depend on OpenStreetMap
  data quality, which varies between cities and regions. Unmapped areas may
  be misclassified.

- **No population data**: the priority map does not include population density.
  High-priority areas with high population density should be considered most
  urgent for intervention. Integrating census data is a recommended next step.

---

# References {.unnumbered}

::: {#refs}
:::

```{r refs-inline, include=FALSE}
# References are declared here for inline citation.
# Full bibliography is rendered above by Quarto/Pandoc.
# The bibliography file (references.bib) is generated by 09_report.R.
```

---

# Technical appendix {.unnumbered}

```{r stats-table}
t(stats) |>
  as.data.frame() |>
  tibble::rownames_to_column("Parameter") |>
  setNames(c("Parameter", "Value")) |>
  kable(caption = "Full summary statistics table",
        align = c("l","l")) |>
  kable_styling(bootstrap_options = c("striped","condensed"),
                font_size = 12, full_width = FALSE)
```

**Software and data sources:**

- R `r R.version$major`.`r R.version$minor` [@rcoreteam]
- terra `r packageVersion("terra")` [@hijmans2022]
- rstac `r packageVersion("rstac")` — STAC API client
- osmdata `r packageVersion("osmdata")` — OSM data
- Landsat Collection 2 Level 2, USGS / Microsoft Planetary Computer [@usgs_c2l2]
- SRTM GL1 DEM, OpenTopography [@farr2007]
- OpenStreetMap contributors, ODbL licence

**Citation for this analysis:**

Richiardi, C., Caroscio, L., Crescini, E., De Marchi, M., De Pieri, G. M.,
Ceresi, C., Baldo, F., Francobaldi, M., & Pappalardo, S. E. (2025). A global
downstream approach to mapping surface urban heat islands using open data and
collaborative technology. *Sustainable Geosciences: People, Planet and
Prosperity*, 100006. <https://doi.org/10.1016/j.susgeo.2025.100006>
